Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-03 Experiments.tex

\begin{figure}[!htb]
        \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak1.pdf}
    \caption{Weak scalability: average pilot duration, average  duration of a
    single AthenaMP execution, and pilot's overhead as a function of different pilot sizes (200, 500, 1000 and 2000 nodes). \jhanote{Caption should be in consistent order as the bars in the figure. Ideally pilot first then AthenMP.}\aanote{DONE}}
\label{fig:weakScal1a}
\end{figure}

>>>>

\begin{figure}[!htb]
        \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak1.pdf}
    \caption{Weak scalability: average pilot duration, average  duration of a
    single AthenaMP execution, and pilot's overhead as a function of different pilot sizes (200, 500, 1000 and 2000 nodes).}
\label{fig:weakScal1a}
\end{figure}

### 2017-04-03 PandaEvolution.tex

The unavailability of pilots is a potential drawback when executing distributed
workloads like MC detector simulation. Pilots are used to increase the
throughput of distributed workloads: while pilots have to wait in the
supercomputer's queue, once scheduled, they can pull and execute jobs
independently from the system's queue. Jobs can be concurrently executed on
every core available to the pilot, and multiple generations of concurrent
executions can be performed until the pilot's walltime is exhausted. This is
particularly relevant for machines like Titan where queue policies privilege
parallel jobs on the base of the number of worker nodes they request: the higher
the number of nodes, the shorter the amount of queue time (modulo fair-share and
allocation policies).

@@

The backfill optimization of Titan's Moab scheduler offers the opportunity to
avoid the overhead of queue wait times without using pilot
abstraction~\cite{maui_backfill_url}. This optimization enables Moab to start
low-priority jobs as long as they do not delay higher priority jobs,
independently on whether the low-priority jobs were queued after the
high-priority jobs.

When the backfill optimization is enabled, users can interrogate Moab about what
number of worker nodes and walltime would be available to a low-priority job at
that moment in time. If a job is immediately submitted to Titan with that number
of worker nodes and walltime, chances are that Moab will immediately scheduled
it, reducing its queue time to a minimum. In this paper, we call this number of
worker nodes and walltime `backfill availability'.

>>>>

The unavailability of pilots is a potential drawback when executing distributed
workloads like MC detector simulation. Pilots are used to increase the
throughput of distributed workloads: while pilots have to wait in the
supercomputer's queue, once scheduled, they can pull and execute jobs
independent from the system's queue. Jobs can be concurrently executed on
every core available to the pilot, and multiple generations of concurrent
executions can be performed until the pilot's walltime is exhausted. This is
particularly relevant for machines like Titan where queue policies privilege
parallel jobs on the base of the number of worker nodes they request: the higher
the number of nodes, the shorter the amount of queue time (modulo fair-share and
allocation policies).

@@

The backfill optimization of Titan's Moab scheduler offers the opportunity to
avoid the overhead of queue wait times without using pilot
abstraction~\cite{maui_backfill_url}. This optimization enables Moab to start
low-priority jobs as long as they do not delay higher priority jobs,
independent of whether the low-priority jobs were queued after the
high-priority jobs.

When the backfill optimization is enabled, users can interrogate Moab about the
number of worker nodes and walltime that would be available to a low-priority job at
that moment in time. If a job is immediately submitted to Titan with that number
of worker nodes and walltime, chances are that Moab will immediately schedule
it, reducing its queue time to a minimum. In this paper, we call this number of
worker nodes and walltime `backfill availability'.

### 2017-04-03 conclusion.tex

More in general, this paper highlighted the fundamental role of WMS for
experimental science. HEP was amongst the first, if not the very first
experimental community to realize the importance of using WMS to manage their
computational campaign(s). As computing become increasingly critical for a range
of other experiments, the impact of WMS on HEP foreshadows the importance of WMS
for other communities.
% experiments and their computational campaigns.
Other experiments such as SKA or LSST, will have their own workload
characteristics, resources types and federation constraints, as well metrics of
performance. The design and experience captured in this paper should prove
invaluable for computational campaigns that need to evaluate the relative merits
of different approaches.

Thanks to the efforts captured in this paper, the time-averaged utilization of
Titan has consistently been upwards of 90\%. Compare this to the time averaged
utilization of other similar leadership machines such as NSF's flagship Blue
Waters, where the average utilization fluctuates between 60-80\% (XDMoD)
machine~\cite{bw-sucks}. These % impressive if not
unprecedented operation gains aside, this work is just a starting point towards
more effective operational models for future leadership and online analytical
platforms\cite{foap-url}, which will have to support ever increasing complex
workloads with varying models for dynamic resource federation. Recent efforts
with Google Cloud for the CMS experiment -- HEPCloud~\cite{hepcloud,google-hep}
represent similar yet different approaches to dynamic resource federation.

Our experience demonstrated also the advantages arising from a systems
engineering and abstractions based approach, in particular the pilot
abstraction. As the type, heterogeneity and complexity of future platforms
(e.g., Exascale, Commercial Clouds) for science increases, there will be a
greater need and emphasis on abstractions to ensure WMS can both utilize as well
as support platforms.

>>>>

More in general, this paper highlighted the fundamental role of WMS for
experimental science. HEP was amongst the first, if not the very first
experimental community to realize the importance of using WMS to manage their
computational campaign(s). As computing become increasingly critical for a range
of experiments, the impact of WMS on HEP foreshadows the importance of WMS for
other experiments (such as SKA, LSST etc.). % and their computational campaigns.
These experiments will have their own workload characteristics, resources types
and federation constraints, as well metrics of performance.  The design and
experience captured in this paper should prove invaluable for designing
computational campaigns and will provide a baseline to evaluate the relative
merits of different approaches.

Thanks to the efforts captured in this paper, the time-averaged utilization of
Titan has consistently been upwards of 90\%. Compare this to the time averaged
utilization of other similar leadership machines such as NSF's flagship Blue
Waters, where the average utilization fluctuates between 60-80\% (XDMoD)
machine~\cite{bw-sucks}. These
% impressive if not
unprecedented efficiency gains aside, this work is just a starting point towards
more effective operational models for future leadership and online analytical
platforms~\cite{foap-url}. These platforms will have to support ever increasing
complex workloads with varying models for dynamic resource federation. Recent
efforts with Google Cloud for the CMS experiment --
HEPCloud~\cite{hepcloud,google- hep} represent similar yet different approaches
to dynamic resource federation.

Our experience demonstrated the advantages arising from a  systems engineering
and abstractions based approach, in particular the pilot abstraction. As the
type, heterogeneity and complexity of future platforms  (e.g., Exascale,
Commercial Clouds) for science increases, there will be a greater need and
emphasis on abstractions to ensure WMS can both utilize as well as support
platforms.
